Double Robust, Flexible Adjustment Methods for Causal Inference: An Overview and an Evaluation

### 2024-07-29 draft/ms-adj-draft.Rmd
replications

---
output:
  # bookdown::word_document2:
    # reference_docx: "word-template.docx"
  bookdown::pdf_document2:
    toc: no
    keep_tex: true
    number_sections: no
    pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
    #latex_engine: xelatex
always_allow_html: true
header-includes:
  #- \usepackage{setspace}\doublespace
  # - \usepackage[nolists, fighead, tabhead]{endfloat}
  # - \usepackage{endnotes}
  # - \let\footnote=\endnote
  # - \setlength{\headheight}{14.5pt}
  # - \setlength{\headheight}{13.6pt}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \lhead{N.I. Hoffmann}
  - \rhead{`r format(Sys.time(), '%B %e, %Y')`}
# - \newcommand{\beginsupplement}{\setcounter{table}{0}  
# \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} 
# \renewcommand{\thefigure}{A\arabic{figure}}}

@@

perform <- function(est_df, label){
  est_df %>%
    # mutate(d = d/truth,
    #        truth = 1) %>%
    summarize(bias = mean(d - truth),
              percent_bias = bias/sd(d),
              rmse = sqrt(mean((ate - truth)^2)),
              mae = median(abs(d - truth))
              # fail_count = first(fail_count),
              # comp_time = first(comp_time)
              ) %>%
    mutate(label = label,
           n = nrow(est_df) + fail_count) %>%
    select(label, everything()) %>%
    return()
}
```

@@

Second, this paper evaluates these methods. They are tested on simulations from @dorie_2019_automated, which cover a range of data-generating processes where ignorability holds -- i.e., there are no unmeasured confounders. They are then used to estimate experimental and observational effects from real-world data, the National Support for Work Demonstration (NSW) originally analyzed by @lalonde_1986_evaluating. In these evaluations, double robust methods are compared to "single robust" methods (i.e., ones with one nuisance model). These include traditional or simpler statistical methods commonly used by social scientists: ordinary least squares (OLS) regression, matching on propensity scores estimated from logistic regression (PSM), and inverse probability weighting (IPW). They are also compared to two more flexible methods that may overcome the misspecification issues that double robust methods aim to overcome: G-computation and the Lin estimator [@lin_2013_agnostic].

@@

Results from the NSW study highlight the importance of ignorability. When observational samples are not comparable to the experimental sample, or when important covariates are not included, all methods -- double robust or traditional -- fail to recover experimental estimates. When appropriate covariates are included, most methods perform well. 
<!-- Exceptions include IPW, DML, and the Lin estimator (surprising given the simulation results), whose estimates tend to be less stable across samples.   -->

@@

# Overview of Techniques  
Each of the methods reviewed in this paper can be thought of as a collection of estimation techniques. Each involves a model for the outcome and another for the treatment exposure, but choice of estimation technique for these two models is left to the discretion of the user. The ways these estimated models relate and are combined into a final effect estimate vary between double robust methods.

@@

More formally, suppose we want to estimate $\tau$ in the following framework:^[Note that this basic DML setup assumes a partially linear model and targets the ATE. If we are interested in the CATE and heterogeneous effects, Chernozhukov et al. [-@chernozhukov_2018_double, p. C35] present an alternative score function that closely resembles AIPW. See also @jacob_2021_cate and @nie_2021_quasioracle.]
$$y_i = \tau d_i + g_0(\mathbf x_i) + u_i,$$
$$d_i = m_0(\mathbf x_i) + v_i.$$

@@

read_csv(here('files/methods.csv')) %>%
  huxtable::hux() %>%
  huxtable::theme_article() %>%
  huxtable::set_width(1) %>%
  # huxtable::set_all_padding(0) %>%
  # huxtable::set_latex_float('h!') %>%
  huxtable::set_caption('Double and single robust methods used for evaluation')

@@

Because many of these methods allow the user to choose the underlying estimation method, results compare three techniques. The first technique uses a logistic regression for the exposure model and an OLS regression for the outcome model. Second is generalized random forests [GRF, @athey_2019_generalized] using the `grf` R package, with separate models for exposure and outcome. The final technique is the SuperLearner (as promoted by the makers of TMLE) using the `SuperLearner` package [@polley_2023_superlearner], again with separate models for exposure and outcome. GLM, glmnet [a weighted average of lasso and ridge regression, @friedman_2021_package], and XGBoost [@chen_2016_xgboost] models are considered for the SuperLearner. These three estimation techniques are used for each of the three double robust methods and for IPW. GRF and SuperLearner are also used for G-computation (OLS predictions with G-computation return identical results to OLS regression coefficient estimates).

For the simulations, results compare only point estimates from these methods. For evaluation of the experimental LaLonde data, I use bootstrapping with 100 samples to obtain standard errors and confidence intervals. Using AIPW as written above results in some wildly biased estimates, due to dividing by some very small propensity scores. Hence I present estimates from a truncated AIPW estimator, where predicted exposure model values are set to 0.01 if they are less than 0.01 and to 0.99 if they are greater than 0.99.

@@

@@

sim_results <- bind_rows(lm_df, psm_df, lin_df, ipw_df, gcomp_df, aipw_trunc_df, tmle_df, dml_df) %>%
  mutate(estimator = factor(estimator, levels = c('ols_logit', 'ols', 'logit', 'grf', 'superlearner')),
         method = factor(method, levels = c('ols', 'psm', 'ipw', 'g-comp', 'lin', 
                                            'aipw', 'tmle', 'dml', 'aipw (original)'))) %>%
  arrange(method, estimator) %>%
  #mutate(estimator = ifelse(is.na(estimator), 'NA', estimator)) %>%
  mutate(bias = ate - truth,
         method_estimator = ifelse(is.na(estimator), as.character(method), paste0(method, ', ', estimator))) %>%
  mutate(method_estimator = factor(method_estimator, levels = unique({.$method_estimator})))

@@

The 20 data sets used here cover a range of these attributes; see the supplemental material from @dorie_2019_automated for details. I use 10 simulations of each data set, resulting in 200 data sets. I then calculate bias, percent bias (the estimator's bias as a percentage of its standard error), root mean squared error (rmse), and median absolute error (mae). I also present the number of datasets for which the method fails and the median computation time for each data set, in seconds.^[Simulations were run on a 2020 MacBook Pro laptop computer with a 2 GHz Quad-Core Intel Core i5 Processor and 16 GB of memory.] In the main text I present average bias and RMSE, while the appendix contains tables with full results (Table \@ref(tab:dorie-results-table)).

@@

Figure \@ref(fig:dorie-linear-bias) shows the bias from each simulation for each method (in the Appendix, Table \@ref(tab:dorie-linear-results) shows the full results). Similarly to the full set of simulations, bias is fairly low for most methods. IPW (logit) and AIPW (OLS/logit) again suffer from greater bias than other methods, though not to quite as extent as in the full range of simulations. Unsurprisingly, methods that assume linearity -- OLS, PSM, the Lin model -- achieve low bias and variance. IPW does less well than expected, even when treatment assignment is modeled with flexible GRF and SuperLearner.

@@

While the AIPW, TMLE, and G-computation with a SuperLearner may be the top performing methods overall, this does not mean that there are some data-generating processes (DGPs) where some other method may do better. Across the the 20 datasets, @dorie_2019_automated vary five DGP characteristics: degree of nonlinearity, the percentage treated, overlap for the treatment group, alignment (correspondence in variables used to generate the exposure and response models), and treatment effect heterogeneity. To test how the methods perform across different values of these characteristics, I begin with the 200 simulations from the 20 datasets used in the main results (Figures \@ref(fig:dorie-results-bias) and \@ref(fig:dorie-results-rmse), Table \@ref(tab:dorie-results-table)). I limit datasets to those generated by a particular value of a DGP characteristic, and I then calculate RMSE for each method for only these datasets. For example, alignment is 0 for datasets 8 and 16 (meaning there is 0 correlation between the terms included in the treatment and outcome models), so for this value of the DGP RMSE is calculated only for those two datasets.

Tables \@ref(tab:dgp) and \@ref(tab:dgp-2) shows the six top-performing methods by lowest RMSE for each each value of the DGP characteristics. Across DGPs, the same methods dominate as in the full range of simulations: the SuperLearner with AIPW, TMLE, and G-compuation. In fact, these three methods take the top three spots for *every* DGP variation in the simulations.

@@

Table \@ref(tab:dorie-size) presents the four methods with the lowest RMSE for each sample size (Table \@ref(tab:dorie-results-size) in the Appendix presents full results).  Again, AIPW, TMLE, and G-computation methods that incorporate the SuperLearner or GRF dominate. There are two exceptions: In the tiny sample of 150, IPW (SuperLearner) and IPW (GRF) figure into the best four methods. PSM is the best-performing method in samples of 1200, but it fails to estimate 17 of the 20 datasets. (With 58 covariates, some of the methods fail in smaller samples.)

@@

size_plot %>%
  # filter(method_estimator %in% best_methods) %>%
  ggplot(aes(x = size, y = rmse, color = method, linetype = estimator, group = method_estimator)) +
  geom_line() +
  ggrepel::geom_text_repel(aes(label = label_min), nudge_x = -1, na.rm = T, 
                           size = text_size, segment.color = 'transparent') +
  ggrepel::geom_text_repel(aes(label = label_max), nudge_x = 1, na.rm = T, 
                           size = text_size, segment.color = 'transparent') +
  scale_y_continuous(trans='log10', breaks = c(0.1, 0.5, 1, 10, 100)) +
  theme(legend.position = 'none') 
```

>>>>

---
output:
  # bookdown::word_document2:
    # reference_docx: "word-template.docx"
  bookdown::pdf_document2:
    toc: no
    keep_tex: true
    number_sections: yes
    pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
    #latex_engine: xelatex
always_allow_html: true
header-includes:
  #- \usepackage{setspace}\doublespace
  # - \usepackage[nolists, fighead, tabhead]{endfloat}
  # - \usepackage{endnotes}
  # - \let\footnote=\endnote
  # - \setlength{\headheight}{14.5pt}
  # - \setlength{\headheight}{13.6pt}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \lhead{N.I. Hoffmann} 
  - \rhead{`r format(Sys.time(), '%B %e, %Y')`}
# - \newcommand{\beginsupplement}{\setcounter{table}{0}  
# \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} 
# \renewcommand{\thefigure}{A\arabic{figure}}}

@@

perform <- function(est_df, label){
  est_df %>%
    # mutate(d = d/truth,
    #        truth = 1) %>%
    summarize(bias = mean(d - truth),
              percent_bias = bias/sd(d),
              rmse = sqrt(mean((ate - truth)^2)),
              mae = median(abs(d - truth))
              # fail_count = first(fail_count),
              # comp_time = first(comp_time)
              ) %>%
    mutate(label = label,
           n = nrow(est_df) + fail_count) %>%
    select(label, everything()) %>%
    return()
}

model_matrix <- function(...){
  options(na.action='na.pass')
  matrix_out <- model.matrix(...)
  options(na.action = 'na.omit')
  return(matrix_out)
}

p_val <- function(estimate, std.error){
  z <- estimate / std.error
  2*(1 - pnorm(abs(as.numeric(z))))
}

compare_tbl <- function(model_list, treatment, tidy = T){
  out_list <- list()
  k <- 1
  for(model_set in model_list){
    names(model_set$grf) <- c('estimate', 'std.error')

    if(tidy == T){
    original_model <- tidy(model_set$original) %>%
      filter(term == treatment)
    } else{
      original_model <- model_set$original
    }

    out_list[[k]] <- original_model %>%
      bind_rows(bind_rows(model_set$grf)) %>%
      bind_rows(model_set$dml) %>%
      bind_rows(model_set$dml_cre) %>%
      bind_rows(model_set$dml_wg_cre) %>%
      mutate(term = c('Original', 'AIPW (GRF)', 'DML (SuperLearner)', 
                      'DML (SuperLearner), CRE FE', 'DML (SuperLearner), hybrid FE'), 
             p.value = p_val(estimate, std.error)) %>%
      remove_rownames() %>%
      tibble()

    k <- k+1
  }
  return(out_list)
}

# compare_tbl_notidy <- function(model_list){
#   out_list <- list()
#   k <- 1
#   for(model_set in model_list){
#     names(model_set[[2]]) <- c('estimate', 'std.error')
#     
#     
#     out_list[[k]] <- model_set[[1]] %>%
#       bind_rows(bind_rows(model_set[[2]])) %>%
#       bind_rows(model_set[[3]]) %>%
#       mutate(term = c('Original', 'AIPW (GRF)', 'DML (SuperLearner)'), p.value = p_val(estimate, std.error)) %>%
#       tibble()
#     
#     k <- k+1
#   }
#   return(out_list)
# }

dml_grf <- function(outcome, treatment, covariates, clustervar = NULL, dataset, 
                    fe = T, target = 'treated', paper, model, drop_na_grf = F, seed = 123){
  set.seed(seed)

  dataset <- filter(dataset, !is.na(!!sym(outcome))) %>%
    select(all_of(c(treatment, outcome, covariates, clustervar)))

  if(drop_na_grf == T){
    dataset <- dataset %>%
      drop_na()
  }

  y <- as.numeric(dataset[[outcome]])
  X <- model_matrix(~., select(dataset, all_of(covariates)))[,]
  d <- as.numeric(dataset[[treatment]])

  grf_model <- grf::causal_forest(X = X, Y = y, W = d, seed = 123,
                        clusters = dataset[[clustervar]])

  grf_out <- grf::average_treatment_effect(grf_model, target.sample = target)

  dataset_drop_na <- as.data.frame(model_matrix(~., dataset)[,]) %>%
    # remove annoying characters
    rename_with(., ~ gsub("'", "", iconv(.x, from = "UTF-8", to='ASCII//TRANSLIT'))) %>%
    rename_with(., ~ gsub("\\[|\\]|\\/|\\*|\\)|\\(", "", .x)) %>%
    rename_with(., ~ gsub(" ", "", .x)) %>%
    # drop missing
    drop_na() %>%
    # remove uninformative columns
    select(where(~n_distinct(.) > 1)) 
  dataset_drop_na <- dataset_drop_na %>%
    left_join(dataset_drop_na %>%
                group_by(!!sym(clustervar)) %>%
                summarize(across(everything(), mean, .names = 'mean_{.col}')) %>%
                ungroup()) %>%
    mutate(intercept = 1)
  # covariates_design <- dataset_drop_na %>% 
  #   select(select(-c(treatment, outcome, clustervar))) %>%
  #   names() %>%
  #   append('intercept', after = 0)
  covariates_design <- names(dataset_drop_na)[!(names(dataset_drop_na) %in% c(treatment, outcome, clustervar)) &
                                                !str_detect(names(dataset_drop_na),  'mean_')]

  # dataset_drop_na$cov_mean <- dataset_drop_na %>%
  #   select(all_of(covariates_design)) %>%
  #   mutate_all(scale) %>%
  #   apply(1, mean)

  # dml_dataframe <- as.data.frame(cbind(y, d, 
  #                                      d_bar = dataset_drop_na[[paste0('mean_', treatment)]],
  #                                      as_tibble(X)[,-1], 
  #                                      as_tibble(mean_X)[,-1], 
  #                                      cluster = dataset_drop_na[[clustervar]]))

  graph_ensemble_regr = gunion(list(
      po("learner", lrn("regr.cv_glmnet", s = "lambda.min")),
      po("learner", lrn('regr.xgboost', max_depth = 4)),
      po("learner", lrn("regr.glm"))
    )) %>>%
      po("regravg", 3)

  ensemble_pipe_regr = as_learner(graph_ensemble_regr)

  # DML
  set.seed(seed)
  dml_data <- double_ml_data_from_data_frame(dataset_drop_na,
                                             x_cols = covariates_design,
                                             y_col = outcome,
                                             d_cols = treatment,
                                             cluster_cols = clustervar)

  # y <- as.numeric(dataset_drop_na[[outcome]])
  # # X <- select(dataset_drop_na, covariates_design)
  # X <- model_matrix(~., select(dataset_drop_na, all_of(covariates)))[,]
  # # mean_X <- model_matrix(~., select(dataset_drop_na, all_of(paste0('mean_', covariates))))[,]
  # d <- as.numeric(dataset_drop_na[[treatment]])
  # 
  # dml_data <- double_ml_data_from_matrix(X = X, y = y, d = d, 
  #                                        cluster_vars = dataset_drop_na[[clustervar]])

  # dml_data <- double_ml_data_from_matrix(X = select(dataset_drop_na, all_of(covariates_design)),
  #                                            y = dataset_drop_na[,outcome],
  #                                            d = dataset_drop_na[,treatment],
  #                                            cluster_vars = dataset_drop_na[,clustervar])
  obj_dml_plr_sim_pipe_ensemble = DoubleMLPLR$new(dml_data,
                                                  ml_l = ensemble_pipe_regr,
                                                  ml_m = ensemble_pipe_regr)
  obj_dml_plr_sim_pipe_ensemble$fit() 
  dml_out <- data.frame(estimate = obj_dml_plr_sim_pipe_ensemble$coef,
                         std.error = obj_dml_plr_sim_pipe_ensemble$se)

  if(fe == T){
    # DML CRE
    set.seed(seed)  
    dml_data_cre <- XTDML::dml_cre_data_from_data_frame(dataset_drop_na,
                                                 x_cols = covariates_design,
                                                 y_col = outcome,
                                                 d_cols = treatment,
                                                 xbar_cols = paste0('mean_', covariates_design[-length(covariates_design)]),
                                                 dbar_cols = paste0('mean_', treatment),
                                                 cluster_cols = clustervar)  
    obj_dml_cre = XTDML::dml_cre_plr$new(dml_data_cre,
                                         ml_l = ensemble_pipe_regr,
                                         ml_m = ensemble_pipe_regr)

    obj_dml_cre$fit()
    dml_cre_out <- data.frame(estimate = obj_dml_cre$coef_theta,
                           std.error = obj_dml_cre$se_theta)

  # DML WG-CRE
  require(parameters)
  set.seed(seed)  
  dml_data_wg_cre <- XTDML::dml_hybrid_data_from_data_frame(dataset_drop_na,
                                               x_cols = covariates_design,
                                               y_col = outcome,
                                               d_cols = treatment,
                                               xbar_cols = paste0('mean_', covariates_design[-length(covariates_design)]),
                                               dbar_cols = paste0('mean_', treatment),
                                               cluster_cols = clustervar)  
  obj_dml_wg_cre = XTDML::dml_hybrid_plr$new(dml_data_wg_cre,
                                       ml_l = ensemble_pipe_regr,
                                       ml_m = ensemble_pipe_regr)

  obj_dml_wg_cre$fit()
  dml_wg_cre_out <- data.frame(estimate = obj_dml_wg_cre$coef_theta,
                         std.error = obj_dml_wg_cre$se_theta)
  } else {
    dml_cre_out <- data.frame(estimate = NA,
                           std.error = NA)
    dml_wg_cre_out <- data.frame(estimate = NA,
                         std.error = NA)
  }

  return(list(grf = grf_out, 
              dml = dml_out,
              dml_cre = dml_cre_out,
              dml_wg_cre = dml_wg_cre_out))
}

@@

Second, this paper evaluates these methods. They are tested on simulations from @dorie_2019_automated, which cover a range of data-generating processes where ignorability holds -- i.e., there are no unmeasured confounders. In the Online Appendix, I also estimate experimental and observational effects from the National Support for Work Demonstration (NSW) originally analyzed by @lalonde_1986_evaluating. In these evaluations, double robust methods are compared to "single robust" methods (i.e., ones with one nuisance model). These include traditional or simpler statistical methods commonly used by social scientists: ordinary least squares (OLS) regression, matching on propensity scores estimated from logistic regression (PSM), and inverse probability weighting (IPW). They are also compared to two more flexible methods that may overcome the misspecification issues that double robust methods aim to overcome: G-computation and the Lin estimator [@lin_2013_agnostic]. Lastly, I use two of the double robust methods to replicate analyses from three recent *American Sociological Review* (ASR) papers that use OLS or logistic regression to estimate causal effects.

@@

Results from the NSW study highlight the importance of ignorability. When observational samples are not comparable to the experimental sample, or when important covariates are not included, all methods -- double robust or traditional -- fail to recover experimental estimates. When appropriate covariates are included, most methods perform well.

In the *ASR* replications, double robust methods provide similar coefficients for most, but not all, of the analyses. In some cases, significance is lost, and in a minority of cases the coefficient sign changes. These replications highlight the importance of testing the robustness of quantitative results by comparing multiple methods and estimators.  
<!-- Exceptions include IPW, DML, and the Lin estimator (surprising given the simulation results), whose estimates tend to be less stable across samples.   -->

@@

# Overview of Techniques  
Each of the methods reviewed in this paper can be used with a variety of estimation techniques, including both traditional statistical methods and flexible machine learning methods. Each involves a model for the outcome and another for the treatment exposure, but choice of estimator for these two models is left to the discretion of the user. Double robust methods are distinct in the ways these estimated models relate and are combined into a final estimate of the desired estimand.

This section provides some intuition for the mathematical theory behind each method and contains R code to simply implement these methods. This is the code used to evaluate these methods later in this paper, and it represents basic implementations of these methods. However, for researchers wishing to put these methods into practice, using a dedicated R packages for each method is probably a better idea. These packages have many flexible options, such as accounting for complex survey design, targeting estimands besides the ATE, and integrating with a variety of estimation techniques. In the Online Appendix, Table \@ref(tab:packages) suggests packages for each method.

@@

More formally, suppose we want to estimate $\tau$ in the following framework:^[Note that this basic DML setup -- presented in the introduction of @chernozhukov_2018_double -- assumes a partially linear model and targets the ATE. If we are interested in the CATE and heterogeneous effects, Chernozhukov et al. [-@chernozhukov_2018_double, p. C35] present an alternative score function that closely resembles AIPW. In this case, DML and AIPW are identical, except DML also includes sample splitting, accounting for the regularization bias that flexible machine learning estimators may induce. See also @jacob_2021_cate and @nie_2021_quasioracle.]
$$y_i = \tau d_i + g_0(\mathbf x_i) + u_i,$$
$$d_i = m_0(\mathbf x_i) + v_i.$$

@@

read_csv(here('files/methods.csv')) %>%
  huxtable::hux() %>%
  huxtable::theme_article() %>%
  huxtable::set_width(1) %>%
  # huxtable::set_all_padding(0) %>%
  # huxtable::set_latex_float('h!') %>%
  huxtable::set_caption('Double and single robust methods used for evaluation. The super learner estimator is an ensemble learning method relying on GLM, glmnet, and XGBoost.')

@@

Because many of these methods allow the user to choose the underlying estimation method, results compare three estimators The first estimator uses a logistic regression for the exposure model and an OLS regression for the outcome model. Second is generalized random forests [GRF, @athey_2019_generalized] using the `grf` R package, with separate models for exposure and outcome. The final estimator is the SuperLearner (as promoted by the makers of TMLE) using the `SuperLearner` package [@polley_2023_superlearner], again with separate models for exposure and outcome. GLM, glmnet [a weighted average of lasso and ridge regression, @friedman_2021_package], and XGBoost with a maximum tree depth of 4 [@chen_2016_xgboost] are the models considered for the SuperLearner. These three estimation techniques are used for each of the three double robust methods and for IPW. GRF and SuperLearner are also used for G-computation (OLS predictions with G-computation return identical results to OLS regression coefficient estimates).

For the simulations, results compare only point estimates from these methods. For evaluation of the experimental LaLonde data, I use bootstrapping with 100 samples to obtain standard errors and confidence intervals. I use the code shown above for AIPW, TMLE, and DML, and I also compare results for two ready-made double robust packages. These include `grf` [@tibshirani_2024_grf], which implements AIPW with generalized random forests, and `DoubleML` [@bach_2024_doubleml], which I use to implement a partially linear DML model with a SuperLearner using GLM, glmnet, and XGBoost.

@@

```{r aipw-package, eval = F}
aipw_package_list <- list()

set.seed(1285)
for(i in 1:length(sims)){
  print(paste0(i, ' out of ', length(sims)))
  dat <- sims[[i]]

  y <- dat$y
  d <- as.numeric(ifelse(dat$z == 'trt', 1, 0))
  x <- model_matrix(~., dat$x)[,]

  grf_model <- grf::causal_forest(X = x, Y = y, W = d)                        
  grf_out <- grf::average_treatment_effect(grf_model, target.sample = 'all', method = 'AIPW', seed = 123)

  aipw_package_list[[i]] <- data.frame(ate = grf_out[['estimate']], 
                              truth = truth, 
                              comp_time = comp_time,
                              dataset = i,
                              dorie_dataset = dat$dataset,
                              set = dat$set,
                              size = dat$size)

  if(i %% 100 == 0){
    write_csv(bind_rows(aipw_package_list), 
          here('files/homemade', 'aipw_package.csv'))
  }

write_csv(bind_rows(aipw_package_list), 
          here('files/homemade', 'aipw_package.csv'))

```{r dml-package, eval = F}
library(DoubleML)
library(mlr3)
library(mlr3pipelines)
library(mlr3learners)
library(mlr3extralearners)

dml_package_list <- list()

set.seed(1285)
for(i in 1:length(sims)){
  print(paste0(i, ' out of ', length(sims)))
  dat <- sims[[i]]

  dml_data <- double_ml_data_from_matrix(X = model_matrix(~., dat$x)[,],
           y = dat$y,
           d = as.numeric(ifelse(dat$z == 'trt', 1, 0)))

  graph_ensemble_regr = gunion(list(
      po("learner", lrn("regr.cv_glmnet", s = "lambda.min")),
      po("learner", lrn('regr.xgboost', max_depth = 4)),
      po("learner", lrn("regr.glm"))
    )) %>>%
      po("regravg", 3)

  ensemble_pipe_regr = as_learner(graph_ensemble_regr)
  set.seed(123)
  obj_dml_plr_sim_pipe_ensemble = DoubleMLPLR$new(dml_data,
                                                  ml_l = ensemble_pipe_regr,
                                                  ml_m = ensemble_pipe_regr)
  obj_dml_plr_sim_pipe_ensemble$fit()

  dml_package_list[[i]] <- data.frame(ate = as.vector(obj_dml_plr_sim_pipe_ensemble$coef), 
                              truth = truth, 
                              comp_time = comp_time,
                              dataset = i,
                              dorie_dataset = dat$dataset,
                              set = dat$set,
                              size = dat$size)

  if(i %% 100 == 0){
    write_csv(bind_rows(dml_package_list), 
          here('files/homemade', 'dml_package.csv'))
  }

write_csv(bind_rows(dml_package_list), 
          here('files/homemade', 'dml_package.csv'))

@@

# aipw_package_df <- read_csv(here('files/homemade', 'aipw_package.csv')) %>%
#   mutate(estimator = 'grf (package)',
#          method = 'aipw')
# 
# dml_package_df <- read_csv(here('files/homemade', 'dml_package.csv')) %>%
#   mutate(estimator = 'superlearner (package)',
#          method = 'dml')

sim_results <- bind_rows(lm_df, psm_df, lin_df, ipw_df, gcomp_df, aipw_trunc_df, tmle_df, dml_df) %>%
  mutate(estimator = factor(estimator, levels = c('ols_logit', 'ols', 'logit', 
          'grf', 'grf (package)', 'superlearner', 'superlearner (package)')),
         method = factor(method, levels = c('ols', 'psm', 'ipw', 'g-comp', 'lin', 
                                            'aipw', 'tmle', 'dml', 'aipw (original)', 'aipw (package)'))) %>%
  arrange(method, estimator) %>%
  #mutate(estimator = ifelse(is.na(estimator), 'NA', estimator)) %>%
  mutate(bias = ate - truth,
         method_estimator = ifelse(is.na(estimator), as.character(method), paste0(method, ', ', estimator))) %>%
  mutate(method_estimator = factor(method_estimator, levels = unique({.$method_estimator})))

@@

The 20 data sets used here cover a range of these attributes; see the supplemental material from @dorie_2019_automated for details. I use 10 simulations of each data set, resulting in 200 data sets. I then calculate bias, percent bias (the estimator's bias as a percentage of its standard error), root mean squared error (rmse), and median absolute error (mae). I also present the number of datasets for which the method fails and the median computation time for each data set, in seconds.^[Simulations were run on a 2020 MacBook Pro laptop computer with a 2 GHz Quad-Core Intel Core i5 Processor and 16 GB of memory.] In the main text I present average bias and RMSE, while the Online Appendix contains tables with full results.

@@

Figure \@ref(fig:dorie-linear-bias) shows the bias from each simulation for each method (table in the Online Appendix). Similarly to the full set of simulations, bias is fairly low for most methods. IPW (logit) and AIPW (OLS/logit) again suffer from greater bias than other methods, though not to quite as extent as in the full range of simulations. Unsurprisingly, methods that assume linearity -- OLS, PSM, the Lin model -- achieve low bias and variance. IPW does less well than expected, even when treatment assignment is modeled with flexible GRF and SuperLearner.

@@

While the AIPW, TMLE, and G-computation with a SuperLearner may be the top performing methods overall, this does not mean that there are some data-generating processes (DGPs) where some other method may do better. Across the the 20 datasets, @dorie_2019_automated vary five DGP characteristics: degree of nonlinearity, the percentage treated, overlap for the treatment group, alignment (correspondence in variables used to generate the exposure and response models), and treatment effect heterogeneity. To test how the methods perform across different values of these characteristics, I begin with the 200 simulations from the 20 datasets used in the main results (Figures \@ref(fig:dorie-results-bias) and \@ref(fig:dorie-results-rmse)). I limit datasets to those generated by a particular value of a DGP characteristic, and I then calculate RMSE for each method for only these datasets. For example, alignment is 0 for datasets 8 and 16 (meaning there is 0 correlation between the terms included in the treatment and outcome models), so for this value of the DGP RMSE is calculated only for those two datasets.

Tables \@ref(tab:dgp) and \@ref(tab:dgp-2) shows the six top-performing methods by lowest RMSE for each each value of the DGP characteristics. Across DGPs, the same methods dominate as in the full range of simulations: the SuperLearner with AIPW, TMLE, and G-computation. In fact, these three methods take the top three spots for *every* DGP variation in the simulations.

@@

Table \@ref(tab:dorie-size) presents the four methods with the lowest RMSE for each sample size (the Online Appendix presents a table with full results). Again, AIPW, TMLE, and G-computation methods that incorporate the SuperLearner or GRF dominate. There are two exceptions: In the tiny sample of 150, IPW (SuperLearner) and IPW (GRF) figure into the best four methods. PSM is the best-performing method in samples of 1200, but it fails to estimate 17 of the 20 datasets. (With 58 covariates, some of the methods fail in smaller samples.)

@@

size_plot %>%
  # filter(method_estimator %in% best_methods) %>%
  ggplot(aes(x = size, y = rmse, color = method, linetype = estimator, group = method_estimator)) +
  geom_line() +
  ggrepel::geom_text_repel(aes(label = label_min), nudge_x = -1, na.rm = T, 
                           size = text_size, segment.color = 'transparent') +
  ggrepel::geom_text_repel(aes(label = label_max), nudge_x = 1, na.rm = T, 
                           size = text_size, segment.color = 'transparent') +
  scale_y_continuous(trans='log10', breaks = c(0.1, 0.5, 1, 10, 100)) +
  theme(legend.position = 'none') +
  labs(x = 'Size (n)')
```

# Replications

```{r aksoy, eval = F}
aksoy <- read_dta(here('replication', 'aksoy.dta')) %>%
  mutate(cid_num = as.integer(factor(cid))) %>%
  as_factor()

# first three models from table 2

m1 <- lm(ISL ~ `_Y28` + `_Y32` + `_Y42` + `_Y46` + `_Y50` + `_Y54` + `_Y57` + `_Y62` + 
     `_Y66` + `_Y70` + `_Y71` + dlen + cid,
   data = filter(aksoy, electyr == T)) # %>%
  # tidy() %>%
  # filter(!str_detect(term, 'cid'), !str_detect(term, '_Y'))

m2 <- lm(ISL ~ `_Y28` + `_Y32` + `_Y42` + `_Y46` + `_Y50` + `_Y54` + `_Y57` + `_Y62` + 
     `_Y66` + `_Y70` + `_Y71` + dlen + cid + gdp_g + t + m + pop,
   data = aksoy) 
  # tidy() %>%
  # filter(!str_detect(term, 'cid'), !str_detect(term, '_Y'))

m3 <- lm(ISL ~ `_Y28` + `_Y32` + `_Y42` + `_Y46` + `_Y50` + `_Y54` + `_Y57` + `_Y62` + 
     `_Y66` + `_Y70` + `_Y71` + dlen + cid + gdp_g + t + m + pop + l_ISL,
   data = aksoy) 
  # tidy() %>%
  # filter(!str_detect(term, 'cid'), !str_detect(term, '_Y'))

aksoy_drop_na_y <- filter(aksoy, !is.na(ISL))

m1_grf_dml <- dml_grf(outcome = 'ISL', 
        treatment = 'dlen', 
        covariates = c('year'), 
        clustervar = 'cid_num',
        dataset = aksoy,
        target = 'all')

m2_grf_dml <- dml_grf(outcome = 'ISL', 
        treatment = 'dlen', 
        covariates = c('year', 'gdp_g', 't', 'm', 'pop'), 
        clustervar = 'cid_num',
        dataset = aksoy,
        target = 'all')

m3_grf_dml <- dml_grf(outcome = 'ISL', 
        treatment = 'dlen', 
        covariates = c('year', 'gdp_g', 't', 'm', 'pop', 'l_ISL'), 
        clustervar = 'cid_num',
        dataset = aksoy,
        target = 'all')

m1 <- data.frame(estimate = 7.159, std.error = 2.539)
m2 <- data.frame(estimate = 7.349, std.error = 2.491)
m3 <- data.frame(estimate = 5.317, std.error = 1.855)

compare_tbl(list(append(list(original = m1), m1_grf_dml), 
                                  append(list(original = m2), m2_grf_dml), 
                                  append(list(original = m3), m3_grf_dml)),
                             tidy = F) %>%
  write_rds(here('replication', 'aksoy_rep.RDS'))

# huxtable::huxreg(compare_tbl(list(append(list(original = m1), m1_grf_dml), 
#                                   append(list(original = m2), m2_grf_dml), 
#                                   append(list(original = m3), m3_grf_dml)),
#                              tidy = F),
#                  statistics = NA) %>%
#   huxtable::insert_row(c('Covariates', 'No', rep('Yes', 2)), after = nrow(.) - 1) %>%
#   huxtable::insert_row(c('Lagged dependent variable', 'No', 'No', 'Yes'), after = nrow(.) - 1) %>%
#   huxtable::set_caption('Replication of aksoy et al. (2022) Table 2, models for outcome of Islamic Votes: "Effect of Fasting Hours (Daylength) during Ramadan on Various Outcome Variables Based on Regression Models That Include Fixed Effects for Provinces and Election Years"') %>%
#   write_rds(here('replication', 'aksoy_tab.RDS'))
# 
# read_rds(here('replication', 'aksoy_rep.RDS')) %>%
#   huxtable::huxreg()

```{r biegert, eval = F}
# Table 3

biegert <- read_dta(here('replication', 'biegert.dta')) %>%
  filter(sampleC == 1) %>%
  mutate(year_num = as.numeric(year),
         allstar_y_num = as.numeric(allstar_y),
         allstar_x_num = as.numeric(allstar_x),
    across(c(allstar_y, allstar_x, year), as.factor),
    player_id_num = as.integer(factor(player_id))) %>%
  as_factor()

m1 <- glm(allstar_y ~ allstar_x, 
          data = filter(biegert, sampleC == 1), family = 'binomial') %>%
  margins::margins(variables = 'allstar_x', 
                   vcov = sandwich::vcovCL(., cluster = ~player_id))

m2 <- glm(allstar_y ~ allstar_x + height + pos + age_0 + age_0_2 + Black + time + time_2 + year, 
          data = filter(biegert, sampleC == 1),
          family = binomial(link = 'logit')) %>%
  margins::margins(variables = 'allstar_x', 
                   vcov = sandwich::vcovCL(., cluster = ~player_id))
